|

### 2021-11-12 01-Chp-Background.Rmd
End of day: 39966

In this thesis, preprints are considered an important source of evidence, in contrast to previous reviews on this topic. However, as with many sources of grey literature,[@mahood2014] there are several logistical issues with carrying out systematic searches in preprint repositories. As such, to enable the inclusion of preprints in the systematic review described in Chapter \@ref(sys-rev-heading), a new tool addressing these issues is presented in Chapter \@ref(sys-rev-tools).

@@

All existing evidence, regardless of study design, is first identified by the by the systematic review presented in Chapter \@ref(sys-rev-heading). Risk-of-bias assessment using a domain-based tool is already a recommended part of the systematic review process, but is particularly important to a triangulation exercise. [@Pagen71; @sterne2019a; @mcguinness2018] As such, a core component of the review is a comprehensive domain-based risk-of-bias assessment for all included studies.

@@

I will will attempt to obtain the raw data from relevant primary studies identified by the systematic review in Chapter \@ref(sys-rev-heading). Any data obtained will be combined with that available from the DPUK portal as part of an individual participant data meta-analysis in Chapter \@ref(ipd-analysis-heading), enabling the assessment of the effect of lipids on dementia stratified by key variables such as sex.
<!----------------------------------------------------------------------------->

@@

-   **Chapter \@ref(background-heading):** Background information on dementia and blood lipid levels. This chapter provides an introduction to the topics covered in this thesis to non-subject area experts, and discusses the motivation for the remainder of the thesis.
-   **Chapter \@ref(sys-rev-tools-heading):** This Chapter introduces a new tool, `medrxivr`, which was used to developed to allow for systematic searches of the health-related preprint repositories.
-   **Chapter \@ref(sys-rev-heading):** This Chapter describes a comprehensive systematic review and meta-analysis of all available evidence on the effect of blood lipids, and interventions that modified blood lipids, on dementia outcomes.
-   **Chapter \@ref(cprd-analysis-heading):** This Chapter examines the relationship between lipid-regulating agent use and dementia outcomes in the Clinical Practice Research Datalink, a large primary care electronic health record database.
-   **Chapter \@ref(ipd-heading):** This Chapter describes an individual patient data analysis of several longitudinal cohort studies to describe the relationship between blood serum lipids and dementia outcomes, stratified by important covariates such as sex.
-   **Chapter \@ref(discussion-heading)**: This Chapter integrates the diverse evidence identified by, and produced as part of, this thesis. The overall strengths and weaknesses of this project are discussed in detail, and further avenues of research are suggested.

An overview of how all research studies in this thesis can be found in Table \@ref(tab:thesisOverview-table)

\blandscape
<!----------------------------------------------------------------------------->
(ref:thesisOverview-caption) Summary of studies included in this thesis, and used as evidence sources in the triangulation exercise performed in Chapter \@ref(tri-heading). Note, Chapter \@ref(sys-rev-tools-heading) is intentionally not included in this table, as it describes a tool rather than a research study.

(ref:thesisOverview-scaption) Summary of studies included in this thesis

```{r thesisOverview-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->
\elandscape

@@

The experience of extracting data for the systematic review in Chapter \@ref(sys-rev-heading) inspired a practical guide for researchers. This piece was co-written with Dr. Emily Hennessy (see Author Declarations in the front materials). <!-- TODO Need to create author declarations section -->

@@

The tool used to visualise the risk-of-bias assessments in Chapter \@ref(sys-rev-heading) has been published in Research Synthesis Methods. See Appendix \@ref(appendix-robvis) for more details on this tool.

>>>>

In this thesis, preprints are considered an important source of evidence, in contrast to previous reviews on this topic. However, as with many sources of grey literature,[@mahood2014] there are several logistical issues with carrying out systematic searches in preprint repositories. As such, to enable the inclusion of preprints in the systematic review described in Chapters \@ref(sys-rev-methods-heading), a new tool addressing these issues is presented in Chapter \@ref(sys-rev-tools).

@@

All existing evidence, regardless of study design, is first identified by the by the systematic review presented inChapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). Risk-of-bias assessment using a domain-based tool is already a recommended part of the systematic review process, but is particularly important to a triangulation exercise. [@Pagen71; @sterne2019a; @mcguinness2018] As such, a core component of the review is a comprehensive domain-based risk-of-bias assessment for all included studies.

@@

I will will attempt to obtain the raw data from relevant primary studies identified by the systematic review in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). Any data obtained will be combined with that available from the DPUK portal as part of an individual participant data meta-analysis in Chapter \@ref(ipd-analysis-heading), enabling the assessment of the effect of lipids on dementia stratified by key variables such as sex.
<!----------------------------------------------------------------------------->

@@

-   **Chapter \@ref(background-heading):** Background information on dementia and blood lipid levels. This chapter provides an introduction to the topics covered in this thesis to non-subject area experts, and discusses the motivation for the remainder of the thesis.
-   **Chapter \@ref(sys-rev-tools-heading):** This Chapter introduces a new tool, `medrxivr`, which was used to developed to allow for systematic searches of the health-related preprint repositories.
-   **Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading):** These Chapters describe, respectively, the methods and results of a comprehensive systematic review and meta-analysis. This review examined all available evidence on the effect of blood lipids, and interventions that modified blood lipids, on dementia outcomes.
-   **Chapter \@ref(cprd-analysis-heading):** This Chapter examines the relationship between lipid-regulating agent use and dementia outcomes in the Clinical Practice Research Datalink, a large primary care electronic health record database.
-   **Chapter \@ref(ipd-heading):** This Chapter describes an individual patient data analysis of several longitudinal cohort studies to describe the relationship between blood serum lipids and dementia outcomes, stratified by important covariates such as sex.
-   **Chapter \@ref(discussion-heading)**: This Chapter integrates the diverse evidence identified by, and produced as part of, this thesis. The overall strengths and weaknesses of this project are discussed in detail, and further avenues of research are suggested.

@@

The experience of extracting data for the systematic review in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading) inspired a practical guide for researchers. This piece was co-written with Dr. Emily Hennessy (see Author Declarations in the front materials). <!-- TODO Need to create author declarations section -->

@@

The tool used to visualise the risk-of-bias assessments in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading) has been published in Research Synthesis Methods. See Appendix \@ref(appendix-robvis) for more details on this tool. Note that this publication does not describe the recently-developed functionality for producing bias direction plots, as described in Chapter \@ref(tri-heading).

### 2021-11-12 03-Chp-Systematic-Review-Methods.Rmd
End of day: 39966

<!-- ## To Do -->

@@

At present, no formalised risk-of-bias assessment tool for Mendelian randomization studies is available. Assessment of the risk of bias in Mendelian randomisation studies was informed by the approach used in a previous systematic review of Mendelian randomisation,[@mamluk2020] as identified by a review of risk-of-bias assessments in systematic reviews of Mendelian randomisation studies (advance results from this review were obtained from contact with the review authors.). <!-- TODO CITATION NEEDED - add once published as a preprint --> A copy of this tool is available in Appendix \@ref(appendix-mr-rob), but in summary, results were assessed for bias arising from weak instruments, genetic and other confounding, pleiotropy and population stratification. Acceptable judgements for each fo the 5 domains in the tool included "low", "moderate" and "high" risk of bias.

>>>>

<!-- TODO Consider reordering the forest plots so that they are the other way around -->

@@

At present, no formalised risk-of-bias assessment tool for Mendelian randomization studies is available. Assessment of the risk of bias in Mendelian randomisation studies was informed by the approach used in a previous systematic review of Mendelian randomisation,[@mamluk2020] as identified by a review of risk-of-bias assessments in systematic reviews of Mendelian randomisation studies (advance results from this review were obtained from contact with the review authors).[@spiga2021] A copy of this tool is available in Appendix \@ref(appendix-mr-rob), but in summary, results were assessed for bias arising from weak instruments, genetic and other confounding, pleiotropy and population stratification. Acceptable judgements for each fo the 5 domains in the tool included "low", "moderate" and "high" risk of bias.

### 2021-11-12 07-Chp-Triangulation.Rmd
End of day: 39966

# Aetiological triangulation across evidence sources {#tri-heading}

@@

## Lay Summary

@@

## Methods
### Data sources

As outlined in the thesis overview presented in Section \@ref(background-thesis-overview), the triangulation draws on the research produced in the preceding chapters.

This chapter builds on the comprehensive systematic review presented in CHapter \@re(sys-rev-heading), and incorporates the results of the analyses presented in Chapters \@ref(cprd-heading) (statins on dementia outcomes in the CPRD) & \@ref(ipd-heading) (lipids on dementia outcomes in IPD).

Table \@ref(triSummary) illustrates the causal question each approach attempted to answer, along with the exposures and outcomes for each.

<!----------------------------------------------------------------------------->
(ref:triSummary-caption) triSummary

(ref:triSummary-scaption) triSummary

```{r triSummary-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

@@

As a first step to assessing the evidence, all evidence sources were grouped by outcome and compared and contrasted. Potential reasons for heterogeneity were examined with specific reference to the risk of bias assessments performed.

This analysis built on the detailed risk of bias assessments performed and presented as part of the systematic review (Chapter \@ref()/\@ref()), and the results of the

@@

In addition to the qualitative discussion of evidence, I attempted to integrate the numerical results of the multiple approaches.

A framework for bias-adjusted meta-analysis

The code supporting the original implementation (written as a STATA do file) was converted into R code, and

This approach incorporated advancements in the way that bias in results is assessed to illustrate both how causal questions could be addressed under this quantitative triangulation framework, and

This was implemented via a bias-adjusted meta-analytical framework.

Several steps were used to accomplish this approach, as described in the following sections. As an illustrative example, the approach was implemented first for a single result and then for all results related to two causal questions of interest, as defined in the following section.

@@

#### Definition of the causal questions of interest (case-studies)
<!-- TODO See ROBINS-E for this, as to how to think about exposures, adn the "Duration and timing of exposure being assessed with different approachesDuration and timing of exposure being assessed with different approaches" section of the aetiological triangulation -->

In the first instance, following standard triangulation

This chapter will consider two questions

Firstly, an ter

Secondly an

Background to the causal question

Following best practice guidance, three

Table with columns describing exposure, outcome, time-point, other aspects (cumulative), and

@@

*	Effect of average exposure to LDL-c during midlife on AD 
*	Effect of cumulative LDL-c from birth to mid-life on AD
* Effect of late-life LDL-c on all-cause dementia

@@

<!----------------------------------------------------------------------------->
(ref:questionsOfInterest-caption) questionsOfInterest

(ref:questionsOfInterest-scaption) questionsOfInterest

```{r questionsOfInterest-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

a)	Proposed method for this chapter
i)	Example: LDL-c at midlife on Alzheimer’s disease risk
ii)	Specific causal effect of interest using ROBINS-E framework:
•	Population of interest: General population
•	Exposure of interest: Low density lipoprotein cholesterol
•	Exposure window of interest: Midlife (45-60)
•	How exposure over time should be summarized: Cumulative exposure
iii)	Identify relevant studies from evidence base
•	LDL-c assessed at midlife
•	MR of LDL-c
•	Statin use at midlife
iv)	Assess risk of bias in each result, define expected directions of each bias and plot

v)	Assign additive value to each level of bias (High = 0.2, Some concerns =0.1, Low = 0), assign sign to these values based on predicted direction and sum across result
vi)	E.g. for effect estimate 0.71 (0.57-0.89):

Element	D1	D2	D3	D4	D5	D6	D7	Overall
Bias	High	Some concerns	Low	Low	Low	Some concerns	Some concerns	Serious
Direction	Left	Away from null	-	-	-	Towards null	Away from null	
Modifier	-0.2	-0.1	0	0	0	0.1	-0.1	-0.3

vii)	Assign a proportional value representing the indirectness for each result
•	Statin studies would give smaller proportion (e.g. 0.7) of effect observed for generally lower lipids, as they would reduce cumulative exposure 
•	SNPs predict quite strongly LDL-c and so would have minimal indirectness
viii)	Combine in a bias-adjusted meta-analysis

@@

@@

####	Assess risk of bias in each result, define expected directions of each bias

Within our framework and as part of the risk-of-bias assessments reported in \@ref(risk-of-bias-subheading), I attempted to records the direction of the bias, so that it could feed into the triangulation.

All of the risk-of-bias tools

In addition, the tool also aim to capture the potential direction of bias for each result. Possible responses included: "Favours experimental", "Favours comparator", "Towards null", "Away from null", and "Unpredictable". Highlight that this is slightly different for the confounding domain in non-randomised studies, which only allows for the "Favours experimental", "Favours comparator", and "Unpredictable" options..

These levels only apply to existing tools, not the MR tool. However, a similar approach was employed.

In line with best practice for meta-analyses, observational studies deemed to be at critical risk of bias were excluded from all quantiative syntheses.

#### Mapping between different risk of bias tools

For the sake of consistency, I mapped the different acceptable judgements in each risk-of-bias tool to a harmonised set: "Low", "Moderate","High". The mapping can be seen in Table \@ref(tab:robLevelsMapping-caption).

Of note, no mapping was performed for the critical risk-of-bias levels present in the tools for non-randomised studies of interventions. This is because current best practice in evidence synthesis is to exclude all studies at critical risk of bias for further quantitative synthesis.

@@

#### Visually inspect results and bias

In order to aid with the triangulation exercise, I created a new method of presentation of these results was developed to enable detailed comparison across different studies contributing to the causal question.

Similar to the standard paired forest plots introduced in Chapter \@ref(), the level of bias across the domains in each study is reported using coloured blocks. However, in contrast to the previous plots, arrows are used in place to illustrate the predicted direction of bias in that domain, categorised as pulling towards the left or the right.

This second aspect required some additional consideration, given the distinct and competing options available to assessors. For example, the direction of the arrow when the direction of bias is thought to favour the intervention/exposure will be be same regardless of the position of the point estimate. The same is not true for the "Towards the null"/"Away from the null" options. For example, if the effect estimate represents a protective effect (below the null), then bias towards the null would be represented by a right pointing arrow (Figure \@ref(fig:)). In contrast, if the effect of the intervention is harmful (effect estimate above the null), bias towards the null will be represented by a right-pointing arrow. The program to create these figures thus requires knowledge of both the predicted direction of bias and the effect estimate.

@@

Helper functions in the forthcoming `triangulate` package - see Appendix BLAH.

@@

These graphs were built using the risk of bias tool described in Appendix \ref()

<!-- TODO Does immortal time bias not also have a consistent effect -->

#### Types of bias

Additive biases are those independent of the effect estimate - they shift it left or right. These biases are indicated by the "Favours comparator"/"Favours experiment/exposure" options in the

@@

For example, for a "High" additive bias in $j$th domain of the $i$th study:

$\delta_{ij}^{\mathrm{High}}$

Uncertainty around this estimate, $\delta_{ij}^{\mathrm{High}}$, is given by a distribution:

$$
\delta_{ij}^{\mathrm{High}} = f(\mu_{i,j}^{\mathrm{High}}, \sigma_{i,j}^{\mathrm{High}})
$$

@@

One key feature of the domain-based risk-of-bias domains is that the domains are considered interchangeable - i.e. a "high" bias in one domain is equivalent to a "high" bias in any other. As such, for this analysis, I assumed that the distributions assigned to each level of bias is contrast across all values of j, or alternatively, is constant across all domains in the risk of bias assessment tool.

@@

<!-- TODO Discuss proportional versus additive, and the fact that all biases considered previously for internal were deemed to be additive. -->

@@

Once the total internal and external additive and proportional biases for each result were computed, they were used to define an adjusted estimate and

Using the method defined in Turner _et al_,[@turner2009] the adjusted estimate is defined as:

@@

The standard error of this estimate is then calculated as:

$$
\mathrm{SE}(\hat{\theta})=\left(\frac{1}{\mu_{i \beta}^{\mathrm{I}} \mu_{i \beta}^{\mathrm{E}}}\right)^{2}\left\{s_{i}^{2}+\left(\mu_{i \beta}^{\mathrm{I}} 2+\sigma_{i \beta}^{\mathrm{I}}\right)\left(\hat{\theta}^{2} \sigma_{i \beta}^{\mathrm{E} 2}+\sigma_{i \delta}^{\mathrm{E} 2}\right)+\sigma_{i \beta}^{\mathrm{I}}\left(\hat{\theta} \mu_{i \beta}^{\mathrm{E}}+\mu_{i \delta}^{\mathrm{E}}\right)^{2}+\sigma_{i \delta}^{\mathrm{I} 2}\right\}
$$

Once each individual result has been adjusted, they are combined n a standard random-effects meta-analysis, conducted using the `metafor` R package.

The code to implement this model was obtained via contact with the primary author of the original paper proposing the elicitation. For compatibility, the STATA code was converted to R, and to verify the accuracy of the translation, the results of the original analyses were reproduced (see Appendix \@ref() for details).

@@

@@

@@

@@

In previous attempts at bias/indirectness-adjusted meta-analysis, the extent to each was assessed via a elicitation process using a number of experts.

@@

i)	Ideally modifying values (both internal and external) would be based on empirical research, but in absence of this, sufficient to model multiple values (High = 0.1, Some concerns =0.05, Low = 0) and relationships (High = 0.3, Some concerns =0.1, Low = 0)?

Need range around modifying values, as total variance of the biases is added to the total variance of the studies (elicitation on scale presented above asks for range rather than single estimate

@@

@@

>>>>

# Aetiological triangulation across new and existing evidence sources {#tri-heading}

@@

## Lay summary

@@

## Methods
### Data sources {#tri-data-sources}

As outlined in the thesis overview presented in Section \@ref(background-thesis-overview), this triangulation exercise draws on the research produced in each of the preceding chapters. More specifically, this chapter builds on the comprehensive systematic review presented in chapter \@re(sys-rev-heading), and incorporates the results of the analyses presented in Chapters \@ref(cprd-heading) (assocation of statins with dementia outcomes in the CPRD) & \@ref(ipd-heading) (association of lipids with dementia outcomes in IPD).

Table \@ref(tab:thesisOverview-table) illustrates the causal question each approach has attempted to address, along with the exposures and outcomes for each.

\blandscape
<!----------------------------------------------------------------------------->
(ref:thesisOverview-caption) Summary of studies included in this thesis, and used as evidence sources in the triangulation exercise. Note, Chapter \@ref(sys-rev-tools-heading) is intentionally not included in this table, as it describes a tool rather than a research study.

(ref:thesisOverview-scaption) Summary of research designs included in this thesis

```{r thesisOverview-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->
\elandscape

@@

As a first step to triangulating the available evidence, all information sources were grouped by outcome, and systematically compared and contrasted. Potential reasons for heterogeneity between study designs were examined with specific reference to the risk-of-bias assessments performed.

<!-- TODO Use the fact that it doesn't make sense to qualitatively compare at the single result level to highlight the need for a more quantitative approach - usually default to summary at the -->

<!-- IDEA Note the text by Turner on the fact that you can just adjust for a single type of bias - a good example of this would be to try and adjust for internal biases so you get the best possible estimate and then apply it to something like a assumed life-course model - steal from NIHR text, and refer to the appendix  -->

This analysis built on the detailed risk of bias assessments performed and presented as part of the systematic review (Chapter \@ref()/\@ref()). Similarly, risk-of-bias assessments were performed for each new source of evidence presented in this thesis, details of which are available in Appendix \@ref(?).

<!-- TODO Need to actually do these, and include results in the analysis. -->

@@

In addition to the qualitative discussion of evidence, I attempted to integrate the numerical results of the multiple approaches. This proof-of-concept approach incorporated advancements in the way that bias in results is assessed (domain-based assessment tools) and existing methods for bias-adjusted meta-analysis[@turner2009] to to illustrate how causal questions could be addressed under this quantitative triangulation framework.

Several steps were used to accomplish this approach, as described in the following sections. In summary these steps are:

1. Define the causal question of interest
1. Identify of relevant evidence sources
1. Assess(and visualise) the extent and direction of bias in each result
1. Define modifying terms for bias and indirectness in each result
1. Perform bias-adjusted meta-analysis using bias-/indirectness-adjusted results

The approach was implemented for two causal questions of interest, as defined below.

@@

#### Definition of the causal questions of interest (case-studies)

<!-- TODO See ROBINS-E for this, as to how to think about exposures, adn the "Duration and timing of exposure being assessed with different approachesDuration and timing of exposure being assessed with different approaches" section of the aetiological triangulation -->

@@

*	Effect of average exposure to LDL-c during midlife on AD 
* Effect of late-life LDL-c on all-cause dementia

@@

@@

individual results were obtained from the general evidence sources described in Section \@ref(tri-data-sources)

In order to illustrate the bias-/indirectness adjustment process detailed in the subsequent sections, the process for calculating the adjusted estimate for a single study is described in detail for a single study. Following this, the process was applied to the two subsets of studies related to the causal questions of interest.

@@

####	Assess risk and direction of bias in each result

The risk of bias tools used to assess each result are described in detail in Section \@ref(risk-of-bias). In summary, randomized controlled trials were assessed using the RoB 2 tool;[@sterne2019] non-randomised studies of interventions (NRSI), were assessed using the ROBINS-I tool;[@sterne2016]

For the sake of consistency, I mapped the different acceptable judgements in each risk-of-bias tool to a harmonised set of judgements: "Low", "Moderate","High". The exact mapping can be seen in Table \@ref(tab:robLevelsMapping-table). Of note, no mapping was performed for the critical risk-of-bias levels present in the tools for non-randomised studies. This is because current best practice in evidence synthesis is to exclude all studies at critical risk of bias for further quantitative synthesis. <!-- TODO CITATION NEEDED -->

@@

In addition to assessing the extent of bias in each domain, as defined by the three levels detailed above, I attempted to records the direction of the bias, so that it could feed into the triangulation.

This was achieved via an additional question at the end of each bias domain, which asked assessors to predict the expected direction of bias in that domain. Note that this question is only apply to existing tools (ROB2/ROBINS-I/ROBINS-E), but I employed an identical approach during assessment of Mendelian randomisation studies. This optional question response were:

* "Favours experimental"/"Favours comparator" (additive bias)
* "Towards null"/"Away from null" (proportional bias)
* "Unpredictable"

As indicated in the options above, the response t this prompt is used both to define the direction and type (additive or proportional) of bias.

Additive biases (identified by a "Favours experimental"/"Favours comparator" response) are XXXX. A key example of additive bias is immortal time bias, which will always favour the intervention.

Proportional biases (identified by a "Towards null"/"Away from null" response) are those related to the magnitude of the effect. A key example of this type of bias is non-differential misclassification bias, which will bias the effect estimate towards the null.

When predicting the direction of effect

From these response the direction of effect was calculated. For additive biases, this was simply a case of shifting the effect estimate to the left or right. For proportional biases, decide

For example, the direction of the bias when the direction of bias is thought to favour the intervention/exposure will be be same regardless of the position of the point estimate. The same is not true for the "Towards the null"/"Away from the null" options.

For example, if the effect estimate represents a protective effect (below the null), then bias towards the null would be represented by a right pointing arrow (Figure \@ref(fig:)). In contrast, if the effect of the intervention is harmful (effect estimate above the null), bias towards the null will be represented by a right-pointing arrow. The program to create these figures thus requires knowledge of both the predicted direction of bias and the effect estimate.

However,

S

Highlight that this is slightly different for the confounding domain in non-randomised studies, which only allows for the "Favours experimental", "Favours comparator", and "Unpredictable" options.

<!-- TODO text -->

#### Visualisation of extent and direction of bias

To aid with the triangulation exercise and to act as, I created a new method of presentation of the extent/direction bias asessment results to enable detailed comparison across different studies contributing to the synthesis. While these figures do not

Similar to the standard paired forest plots introduced in Chapter \@ref(sys-rev-), the level of bias across the domains in each study is reported using coloured blocks. However, in contrast to the previous plots, arrows are used in place to illustrate the predicted direction of bias in that domain, categorised as pulling towards the left or the right.

This second aspect required some additional consideration, given the distinct and competing options available to assessors.

@@

Helper functions in the forthcoming `triangulate` package - see Appendix \@ref() - help to preprocess risk of bias data for use with both bias-direction plots and adjusted meta-analyses. These bias direction plots themselves were created using the risk of bias isualisation tool described in Appendix \ref().

@@

<!-- TODO Does immortal time bias not also have a consistent effect? yes, but inception bias does not necessarily -->

@@

For example, for a "High" additive bias in $j$th domain of the $i$th study is deifined as $\delta_{i,j}^{\mathrm{High}}$ and uncertainty around this estimate is given by a distribution:

$$
\delta_{i,j}^{\mathrm{High}} = f(\mu_{i,j}^{\mathrm{High}}, \sigma_{i,j}^{\mathrm{High}})
$$

@@

One key feature of the domain-based risk-of-bias domains is that the domains are considered interchangeable - i.e. a "high" bias in one domain is equivalent to a "high" bias in any other. As such, for this analysis, I assumed that the distributions assigned to each level of bias are identical across all values of $j$, that is, are identical for all domains in the risk of bias assessment tool.

@@

(ref:priorsAdd-caption) Distributions for additive bias assigned to each bias judgement under two sensitivity scenarios, on the log scale.

(ref:priorsAdd-scaption) Additive bias distributions

```{r priorsAdd-table, message=FALSE, results="asis", echo = FALSE}
```

@@

Once the total internal and external additive and proportional biases for each result were computed, they were used to define an adjusted estimate for each result included in the synthesis

Using the method defined in Turner _et al_,[@turner2009] the adjusted estimate for each result is defined as:

@@

where $\mu_{i\delta}$ denotes total additive and $\mu_{i\beta}$ the total proportional bias, and the $^{\mathrm{I}}$ and $^{\mathrm{E}}$ superscripts denote internal and external (indirectness) biases respectively.

using the same terms, the standard error of this estimate, $\hat{\theta}$, is then calculated as:

$$
\mathrm{SE}(\hat{\theta})=\left(\frac{1}{\mu_{i \beta}^{\mathrm{I}} \mu_{i \beta}^{\mathrm{E}}}\right)^{2}\left(s_{i}^{2}+\left(\mu_{i \beta}^{\mathrm{I}} 2+\sigma_{i \beta}^{\mathrm{I}}\right)\left(\hat{\theta}^{2} \sigma_{i \beta}^{\mathrm{E} 2}+\sigma_{i \delta}^{\mathrm{E} 2}\right)+\sigma_{i \beta}^{\mathrm{I}}\left(\hat{\theta} \mu_{i \beta}^{\mathrm{E}}+\mu_{i \delta}^{\mathrm{E}}\right)^{2}+\sigma_{i \delta}^{\mathrm{I} 2}\right)
$$

Once each individual result has been adjusted, they are combined n a standard random-effects meta-analysis <!-- TODO CITATION NEEDED -->

#### Software

Personal communication with the authors of the original method resulted in the STATA code to implement the

As detailed in the introduction to this section, the model for defining bias-indirectness-adjusted results was available as STATA code. The  has since been generalised as part of the `triangulate` R package to allow for users to apply the approach detailed here.

The package allows for preprocessing of domain-based risk-of-bias data to correctly identify the direction of the bias relative to the effect estimate, calculate bias/indirectness-adjusted estimates for each result, and synthesise these adjusted results in a standard random-effects meta-analysis.

All meta-analyses conducted within the package are implemented using the `metafor` R package.

@@

```{r biasDirectionPlots, include = F}
```

(ref:ldlAdBiasDirection-cap) __Bias direction plot, summarising internal biases related to effect of LDL-c on Alzheimer's disease :__

(ref:ldlAdBiasDirection-scap) Bias direction plot, summarising internal biases related to the

```{r ldlAdBiasDirection, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:ldlAdBiasDirection-cap)', out.width='100%', fig.scap='(ref:ldlAdBiasDirection-scap)'}
knitr::include_graphics(file.path("figures/tri/midlife_AD.png"))
```
&nbsp;<!----------------------------------------------------------------------->

@@

As expected

@@

Main takeway is that heteorgeneity between study designs is substantially reduced

@@

Previous work on assessing the impact of different sources of bias on effect estimates . For example, a study examining the effect of found that studies with

Previous work examining

A further issue is the "Low" risk of bias domain. In this proof-of-concept analysis, I assumed that domains at "Low" risk of bias did not require any adjustment. However, "Low" risk of bias is not equivalent to the absence of bias, and potentially should still be adjusted for. In this case, defining the predicted direction of bias would be challenging in the absence of obvious sources of bias (particularly given how challenging it can be when sources of bias are evident within a domain).

In previous attempts at bias/indirectness-adjusted meta-analysis, the extent to each was assessed via a elicitation process using a number of experts. This approach is potentially subject to misclassification of the impact of bias on the basis of the results, as there is no way to ensure that results at a similar level of bias for confounding (for example) are being adjusted by the same amount. For example, if experts are influenced by the absolute effect estimate, then a result demonstrating a stronger protective effect may receive greater modification than a study at the same "risk" of confounding but with a more modest effect. This is particular true where experts have preconceived ideas about the true effectiveness of the intervention.

In this case, separating the assessment of bias from the assignment of modifying values to each level of bias will limit the potential for. Alternatively, reasonable modifying distributions for each level of bias could be defined _a priori_ by the study team, similar to how important confounders and co-interventions are defined in advance when performing ROBINS-I/ROBINS-E assessments.

In the mean time, the extent of the impact of knowledge of the results could be analysed using existing data. For example, a re-analysis of the data presented in a recent paper where both expert eliciation and structured

@@

@@

This method is also different from previous implementations of the bias-adjusted meta-analysis approach, in that for each domain for each study, the bias can be either proportional or. IN previous implementations, distributions for each source of bias (pre-domain-based tools) were elicted on both the additive and proportional scales.[@turner2009] Puts more pressure on the assessor to correct sum additive and proportional biases highlighted, and answer with only a single option.

@@

### Filling in the gaps

Triangulation as a extension of evidence synthesis.

### 2021-11-12 front-and-back-matter/_00-introduction.Rmd
End of day: 39966

```{r word-count, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)

options(scipen=999)

@@

```{r word-count-fig, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(mapping = aes(x=date,y=words)) +
  geom_line(data = target_df, colour = "black", linetype = 2) +
  geom_hline(yintercept = 40000, colour = "black", linetype = 2) +
  geom_vline(xintercept = as.numeric(as.Date("20211116", format="%Y%m%d")), colour = "black", linetype = 2) +
  geom_vline(xintercept = as.numeric(as.Date("20211216", format="%Y%m%d")), colour = "red") +
  scale_x_date(limits = c(as.Date("20201124", format="%Y%m%d"),as.Date("20211231", format="%Y%m%d"))) +
  geom_step(data = words, colour = "red")

>>>>

```{r wordcount, echo=FALSE, warning=FALSE, message=FALSE}
options(scipen=999)

@@

```{r wordcount-fig, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(mapping = aes(x=date,y=words)) +
  geom_line(data = target_df, colour = "black", linetype = 2) +
  geom_hline(yintercept = 40000, colour = "black", linetype = 2) +
  geom_vline(xintercept = as.numeric(as.Date("20211116", format="%Y%m%d")), colour = "black", linetype = 2) +
  geom_vline(xintercept = as.numeric(as.Date("20211216", format="%Y%m%d")), colour = "red") +
  scale_x_date(limits = c(as.Date("20201124", format="%Y%m%d"),as.Date("20211231", format="%Y%m%d"))) +
  geom_step(data = words, colour = "red") +
  geom_text(aes(x = as.Date("20210910", format="%Y%m%d"), y = 30000, label = "Tables"))
